Secure Composition \\ of Robust and Optimising Compilers \vspace{-1.5em}

### 2024-09-12 csf2/main.tex
Paper changes

In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we have the classic definition of robust preservation~\cite{abate2019jour}.
\begin{definition}[Robust Preservation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  $\;$

@@

Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
% 
Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space.

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
  % Given \texttt{p}, yield
  \lambda \texttt{p}\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

% Examples of this are present in industry:
% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
% Both languages can be used at the same time in one project~\cite{androidstudio}.
% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
Now, the following theorem tells us what happens if these are secure:
Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  $\;$

  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
\end{theorem}

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
\begin{definition}[Lower Composition]
  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
  % Given $L$ and $\src{p}$, yield
  \lambda \src{p}, L\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

% The following theorem demonstrates what happens if the involved compilers are secure:
Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  $\;$

  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
\end{theorem}

@@

Future work intends to use upper and lower composition as means to build a secure compiler incrementally
This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.

@@

\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
Moreover, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).

@@

\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}

$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
If the register is marked inactive, secrets may appear on traces.
The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.

@@

Memory accesses to secret data need to be present to reason about memory safety, even when execution is in data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety. 
Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).

Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.

The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
When the data independent timing mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.

@@

\begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
  In each half, the left column (\textit{Active}) has data-independent timing $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
\end{exampleenv}
\begin{figure}[!htb]
	\vspace{-1em}\small
  $$
  % {
  \begin{array}{ccc}
    & \Lscct &        \\
    \text{Active} & \mid & \text{Inactive} \\\hline
    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \vdots & \mid & \vdots \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
  \end{array}
  % }
  $$
  $$
  % {
  \begin{array}{ccc}
          & \text{Specification} & \\
  \text{Active} & \mid & \text{Inactive} \\\hline
     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \vdots & \mid & \vdots \\
      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
  \end{array}
  % }
  $$
  \caption{Traces for \Cref{ex:lscct}.}
  \label{fig:ex-cct}
\end{figure}

@@

This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.

@@

Our secure compilation proofs rely on established proof techniques called backtranslation~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context. 
All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are essentially an identity function). 
% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.

@@

>>>>

In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
\begin{definition}[Robust Preservation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  $\;$

@@

Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
% 
Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$ as well as for classes of hyperproperties, but these results have been moved to the technical report in the interest of space.

@@

%\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
%Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
%We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
%Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
%\begin{definition}[Upper Composition]
%  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
%  % Given \texttt{p}, yield
%  \lambda \texttt{p}\ldotp
%  \left\{\mbox{\begin{tabular}{c}
%    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
%    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
%  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
%\end{definition}

% Examples of this are present in industry:
% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
% Both languages can be used at the same time in one project~\cite{androidstudio}.
% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
%Now, the following theorem tells us what happens if these are secure:
%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

%\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
%  $\;$
%
%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
%\end{theorem}

%Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
%{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
%Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
%\begin{definition}[Lower Composition]
%  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
%  % Given $L$ and $\src{p}$, yield
%  \lambda \src{p}, L\ldotp
%  \left\{\mbox{\begin{tabular}{c}
%    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
%    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
%  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
%\end{definition}

@@

% The following theorem demonstrates what happens if the involved compilers are secure:
%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

%\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
%  $\;$ 
%
%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
%\end{theorem}

@@

%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.

@@

\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
Moreover, programs that operate on secrets, have been compiled with FaCT~\cite{cauligi2019fact}, and a ``data independent timing mode'' enabled, do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).

@@

\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}

$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
This mode is inspired by a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
If the register is marked inactive, secrets may appear on traces.
The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.

@@

Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety. 
Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).

Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.

The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.

@@

\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.

  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
\end{exampleenv}
\begin{figure}[!htb]
	\vspace{-1em}\small
  $$
  % {
  \begin{array}{ccc}
    & \Lscct &        \\
    \text{Active} & \mid & \text{Inactive} \\\hline
    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \vdots & \mid & \vdots \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
  \end{array}
  % }
  $$
  $$
  % {
  \begin{array}{ccc}
          & \text{Specification} & \\
  \text{Active} & \mid & \text{Inactive} \\\hline
     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \vdots & \mid & \vdots \\
      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
  \end{array}
  % }
  $$
  \caption{Traces for \Cref{ex:lscct}.}
  \label{fig:ex-cct}
\end{figure}

@@

This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
This simple flag combines the effect of FaCT~\cite{cauligi2019fact}

@@

Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context. 
All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.

@@

%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}

There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
Our framework does not consider exploits.
%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
